CONEM2024-0025\\ ABC MULTICRITERIA CLASSIFICATION WITH ANALYTICAL HIERARCHY PROCESS: HOW TO FORCE CONSISTENCY OF PAIRWISE COMPARISONS MATRIX

### 2024-02-02 accepted/CONEM2024/FCPCOM/CONEM2024.tex
updating

The question is: is the difference between the importance of a product with a 30-day lead time and a product with a 7-day lead time really only 0.35\%. This importance can only really be appreciated when put in comparison with the difference in importance of 12\% between the product with the highest turnover and the product with the lowest turnover.

Analyzing these values, we come to the conclusion that it seems coherent to adopt as a measure for the weights not gross values, but comparative values. This conclusion gave rise to the idea of assigning comparative weights.

>>>>

The question is: is the difference between the importance of a product with a 30-day lead time and a product with a 7-day lead time only 0.35\%? This value can only be appreciated when compared with the difference in importance of 12\% between the product with the highest turnover and the product with the lowest turnover. It is like creating an average of the annual sales of vehicles and bananas, hoping that this average will have some meaning. For there to be real meaning in the weighted average found by the multicriteria method, it is necessary to apply some procedure to normalize the weights assigned to each criterion.

	\citet{Odu2019} cites CRITIC as one of the methods applied for weight assignment in multicriteria classification. This method starts with normalizing the weights using the following equations:

	The problem with this normalization method is that the largest values of the different criteria will have the same importance. In the case of the companies studied, this may not correspond to the assessment of experts. However, it is possible to make a comparative analysis of amounts using different criteria. And use this analysis for normalization. For this purpose, we propose the following procedure:

\begin{itemize}
\item classification of products in descending order according to a first criterion;
\item identification of the product with the highest weight, according to a second criterion;
\item peer-to-peer comparison with the products ordered according to the first criterion, answering the following question: the degree of importance of this product according to the second criterion is equivalent to what degree of importance of the other product evaluated according to the first criterion?
\end{itemize}

Using this procedure, we can identify, for example, that the lead time weight of a product that has a lead time of 30 days is equivalent to the annual turnover weight of a product with an annual turnover of 68,380.00 reais.

    \section{REFERENCES}
